TSELM: Target Speaker Extraction using Discrete Tokens and Language Models

### 2024-09-12 src/paper.tex
update

The discretization of audio has gained significant attention with the advancement of language models (LMs) \cite{l1, l2, l3, l4, l5}. It has been explored in
tasks such as speech enhancement \cite{selm} and blind speech separation \cite{dasb,tokensplit}.
This method converts audio into discrete tokens, and leverages LMs to model them, thereby simplifying audio generation tasks by transforming complex regression problems into classification tasks \cite{dasb}. Self-Supervised Learning (SSL) models such as HuBERT \cite{hubert} and WavLM \cite{wavlm} have demonstrated outstanding performances across numerous downstream tasks \cite{superb}, as they extract continuous representations rich in semantic and timbral information from speech.

@@

\section{Conclusion}
In this work, we introduce a novel way using discrete tokens and language models for target speaker extraction.
Our method leverages multiple hidden layers of WavLM and Kmeans tokenizers for encoding, employs cross-attention and a language model for separation, and utilizes a scalable HiFi-GAN for audio reconstruction.
Experiments have 
shown that our model can achieve excellent performance in terms of speech quality, and comparable performance in terms of speech intelligibility. However, we observe a gap between discrete methods and continuous methods especially in speech intelligibility and speaker similarity. Future research should focus on shrinking 
this gap.

>>>>

The discretization of audio has gained significant attention with the advancement of language models (LMs) \cite{l1, l2, l3, l4, l5}. It has achieved superior performances compared with discriminative models in
tasks such as speech enhancement \cite{selm} and blind speech separation \cite{dasb,tokensplit}.
These methods convert audio into discrete tokens, and leverages LMs to model them, thereby simplifying audio generation tasks by transforming complex regression problems into classification tasks \cite{dasb}.

Self-Supervised Learning (SSL) models such as HuBERT \cite{hubert} and WavLM \cite{wavlm} have demonstrated outstanding performances across numerous downstream tasks \cite{superb}, as they extract continuous representations rich in semantic and timbral information from speech.

@@

\section{Conclusion}
In this work, we introduce a novel way using discrete tokens and language models for target speaker extraction.
Our method leverages multiple hidden layers of WavLM and Kmeans tokenizers for encoding, employs cross-attention and a language model for separation, and utilizes a scalable HiFi-GAN for audio reconstruction.
Experiments have 
shown that our model can achieve excellent performance in terms of speech quality, and comparable performance in terms of speech intelligibility. However, we observe a gap between discrete methods and continuous methods especially in speech intelligibility and speaker similarity. We will focus on shrinking 
this gap.
